---
title: "Data Manipulation"
output: github_document
---
```{r setup}
library(tidyverse)
```

## Load the FAS litters data 

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

## 'select'

Choose some columns and not others
```{r}
select(litters_df, group, gd0_weight)
```

: indicates you select a range of columns
```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
```

Specify what we want to get rid of 
```{r}
select(litters_df, -litter_number)
```

Rename columns
```{r}
select(litters_df, GROUP = group, LITTER_NUMBER = litter_number)
```

Can also use rename function to rename a column without losing the other columns
```{r}
rename(litters_df, GROUP = group, LITTER_NUMBER = litter_number)
```

Select helpers - functions that help you select a column based on a criteria
```{r}
select(litters_df, starts_with("gd"))
```

Select to move litter number to the beginning and keep everything
```{r}
select(litters_df, litter_number, everything())
```

does the same thing as above, moves columns to the front
```{r}
relocate(litters_df, litter_number)
```

## ' filter'

Removes or keeps rows that you're interested in retaining for select analyses
Need to give it a variable and a condition so it knows what to do 

```{r}
filter(litters_df, gd0_weight < 22)
```

```{r}
filter(litters_df, gd0_weight >= 22)
```

== test for equality
```{r}
filter(litters_df, gd_of_birth == 20)
```

negate things --> gd_of_birth when it is NOT equal to 20
```{r}
filter(litters_df, !(gd_of_birth == 20))
```

2 conditions
```{r}
filter(litters_df, gd0_weight >= 22, gd_of_birth == 20)
```

Character variables 
```{r}
filter(litters_df, group == "Mod8")
```

2 conditions of characters (either of these 2 options)
```{r}
filter(litters_df, group %in% c("Con7", "Mod8"))
```

## 'mutate'

Modify existing variable or create a new variable 
Modify group variable to making everything lower case
```{r}
mutate(litters_df, 
       wt_gain = gd18_weight - gd0_weight,
       group = str_to_lower(group))
```

## 'arrange'

Put things in order (ascending)
Sort by 2 things - first by the first variable and then again in order within those levels
```{r}
arrange(litters_df, pups_born_alive)
arrange(litters_df, pups_born_alive, gd0_weight)
```


## pipe operation %>%

Not using pipes - many steps
```{r}
litters_data_raw = read_csv("./data/FAS_litters.csv")
litters_clean_name = janitor::clean_names(litters_data_raw)
litters_data_selected = select(litters_clean_name, -pups_survive)
litters_mutated = mutate(litters_data_selected, wt_gain = gd18_weight - gd0_weight)
litters_without_missing = drop_na(litters_mutated, gd0_weight)
```

USE THE PIPE OPERATOR INSTEAD

```{r}
litters_df = 
  read_csv("./data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  select(-pups_survive) %>%
  mutate(wt_gain = gd18_weight - gd0_weight) %>% 
  drop_na(gd0_weight)
```

Both above codes produce the same thing but the pipe is better