---
title: "Data Import"
output: github_document
---
```{r setup}
library(tidyverse)
```

## Read in some data

Read in the litters dataset. 
"read_csv" exisits in the readr package in tidyverse
Since we're working within the "data_wrangling_i" project, we can create a relative path starting with that folder
  "./" means start with the project folder
  
Also will update names in df because they aren't in good form (clean_names)
  Doing "janitor::" means clean_names is in that package. But since    it's the only function in that package we need, we don't need to     load the entire package

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

## Take a look at the data

Printing in the console
```{r}
litters_df
```

Prints first 6 observations
```{r}
head(litters_df)
```

Print last 6 observations 
```{r}
tail(litters_df)
```

Some descriptive info
```{r}
skimr::skim(litters_df)
```

Use view(litter_df) in console to open up a new tab of the data set. Run this in the console because it makes knitting easier
