---
title: "Data Manipulation"
output: github_document
---
```{r setup}
library(tidyverse)
```

## Load the FAS litters data 

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

## 'select'

Choose some columns and not others
```{r}
select(litters_df, group, gd0_weight)
```

: indicates you select a range of columns
```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
```

Specify what we want to get rid of 
```{r}
select(litters_df, -litter_number)
```

Rename columns
```{r}
select(litters_df, GROUP = group, LITTER_NUMBER = litter_number)
```

Can also use rename function to rename a column without losing the other columns
```{r}
rename(litters_df, GROUP = group, LITTER_NUMBER = litter_number)
```

Select helpers - functions that help you select a column based on a criteria
```{r}
select(litters_df, starts_with("gd"))
```

Select to move litter number to the beginning and keep everything
```{r}
select(litters_df, litter_number, everything())
```

does the same thing as above, moves columns to the front
```{r}
relocate(litters_df, litter_number)
```

