---
title: "Tidy Data"
output: github_document
---

```{r setup}
library(tidyverse)
```


## 'pivot_longer'

Load the pulse data (sas format)

```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names()
```

Wide to long format
tell which columns need to be made wider in pivot_longer function
Those column names need to be put into another variable ("names_to")
Values to into ("values_to") bdi
use "name_prefix" to remove the prefix of bdi_score
```{r}
pulse_data_tidy = 
  pulse_data %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi"
  )
```

Put everuthing into a single code chuck
Rewrite, combine and extend (add mutate step)
"recode" exisits in dplyr to modify variables
```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi"
  ) %>% 
  relocate(id, visit) %>%
  mutate(visit = recode(visit, "bl" = "00m"))
```


## 'pivot_wider'

Make up some data
Take values in a column and make new columns out of that
```{r}
analysis_result = 
  tibble(
    group = c("treatment", "treatment", "placebo", "placebo"),
    time = c("pre", "post", "pre", "post"),
    mean = c(4, 8, 3.5, 4)
  )

analysis_result %>% 
  pivot_wider(
    names_from = "time", 
    values_from = "mean"
  )
```


## Binding rows

Using Lord of the rings data
What happens when I have data in multiple tables and I need to stack those rows up?

Step 1: import each table

```{r}
fellowship_ring = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") %>% 
  mutate(movie = "fellowship_ring")

two_towers = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") %>% 
  mutate(movie = "two_towers")


return_king = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") %>% 
  mutate(movie = "return_king")
```

Bind all the rows together (stack up these tables)

```{r}
lotr_tidy = 
  bind_rows(fellowship_ring, two_towers, return_king) %>% 
  janitor::clean_names() %>%
  relocate(movie) %>% 
  pivot_longer(
    female:male,
    names_to = "gender",
    values_to = "words"
  )
```

## Joining Datasets

Import the FAS datasets

Get info from litters dataset into the pups dataset (litter number is the variable that exisits in both df)

Litters DF: Treatment group and treatment day are in the same variable (group), so we want to separate them (split them by the number of characters)

```{r}
pup_df = 
  read_csv("./data/FAS_pups.csv") %>% 
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female"))

litters_df = 
  read_csv("./data/FAS_litters.csv") %>% 
  janitor::clean_names() %>%
  relocate(litter_number) %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3)
```

Now we join them (doing a left join)
Merch liters data into the pups data (pups data is left and liters data is right)
Join by litter number
```{r}
fas_df =
  left_join(pup_df, litters_df, by = "litter_number") %>% 
  arrange(litter_number) %>% 
  relocate(litter_number, dose, day_of_tx)
  
```




